---
title: "week_3_exercise"
author: "Tim Fässler"
date: "5/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation

```{r, warning = FALSE, message = FALSE}
library(readr)
library(sf)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(SimilarityMeasures)

```

# Task 1: Segmentation
```{r}
data <- read_delim("caro60.csv", ",")
caro <- data.frame(data)
  

caro <- caro %>%
  mutate(
    nMinus3 = sqrt((lag(E,3)-E)^2+(lag(N,3)-N)^2),
    nMinus2 = sqrt((lag(E,2)-E)^2+(lag(N,2)-N)^2),
    nMinus1 = sqrt((lag(E,1)-E)^2+(lag(N,1)-N)^2),
    nPlus1  = sqrt((E-lead(E,1))^2+(N-lead(N,1))^2), 
    nPlus2  = sqrt((E-lead(E,2))^2+(N-lead(N,2))^2),
    nPlus3  = sqrt((E-lead(E,3))^2+(N-lead(N,3))^2)
  )

caro <- caro %>%
  rowwise() %>%
  mutate(
    stepMean = mean(c(nMinus3, nMinus2, nMinus1,nPlus1,nPlus2, nPlus3))
  ) %>%
  ungroup()

summary(caro$stepMean)
```

# Task 2: Specify and apply threshold d

```{r}
hist(caro$stepMean)

caro <- caro %>% 
  ungroup() %>%
  mutate(static = stepMean < 10)
```

# Task 3: Visualize segmented trajectories

```{r}

caro_filter <- caro %>%
  filter(!static)

caro_filter%>%
  ggplot(aes(E, N))  +
  geom_path() +
  geom_point() +
  coord_fixed() +
  theme(legend.position = "bottom") +
  ggtitle("Only points while moving")


ggplot() +
  geom_path(caro_filter, mapping = aes(E, N)) +
  geom_point(caro, mapping = aes(E, N, color = static)) +
  ggtitle("Static vs moving points")
```
# Task 4: Segment-based analysis

```{r}
rle_id <- function(vec){
  x <- rle(vec)$lengths
  as.factor(rep(seq_along(x), times=x))
}

caro <- caro %>% mutate(segment_id = rle_id(static))

seg_colors <- filter(caro, static == FALSE)
seg_long <- seg_colors %>% group_by(segment_id) %>% mutate(long = length(segment_id)) %>% filter(long >= 5)

ggplot() +
  geom_path(seg_colors, mapping = aes(E, N, color = seg_colors$segment_id)) +
  geom_point(seg_colors, mapping = aes(E, N, color = seg_colors$segment_id)) +
  scale_color_discrete("Segment Number") +
  ggtitle("All Segments")

ggplot() +
  geom_path(seg_long, mapping = aes(E, N, color = seg_long$segment_id)) +
  geom_point(seg_long, mapping = aes(E, N, color = seg_long$segment_id)) +
  scale_color_discrete("Segment Number") +
  ggtitle("All Segments with >= 5 Nodes")


```

# Task 5

```{r}

pedestrian <- read_delim("pedestrian.csv", ",") %>% data.frame()

tra1 <- ggplot() +
  geom_point(pedestrian, mapping = aes(E, N), alpha = 0.1) +
  geom_point(subset(pedestrian, TrajID %in% "1"), mapping = aes(E, N), color = "red") +
  geom_path(subset(pedestrian, TrajID %in% "1"), mapping = aes(E, N), color = "red") +
  ggtitle("TrajID: 1")

tra2 <- ggplot() +
  geom_point(pedestrian, mapping = aes(E, N), alpha = 0.1) +
  geom_point(subset(pedestrian, TrajID %in% "2"), mapping = aes(E, N), color = "yellow") +
  geom_path(subset(pedestrian, TrajID %in% "2"), mapping = aes(E, N), color = "yellow") +
  ggtitle("TrajID: 2")

tra3 <- ggplot() +
  geom_point(pedestrian, mapping = aes(E, N), alpha = 0.1) +
  geom_point(subset(pedestrian, TrajID %in% "3"), mapping = aes(E, N), color = "green") +
  geom_path(subset(pedestrian, TrajID %in% "3"), mapping = aes(E, N), color = "green") +
  ggtitle("TrajID: 3")

tra4 <- ggplot() +
  geom_point(pedestrian, mapping = aes(E, N), alpha = 0.1) +
  geom_point(subset(pedestrian, TrajID %in% "4"), mapping = aes(E, N), color = "darkgreen") +
  geom_path(subset(pedestrian, TrajID %in% "4"), mapping = aes(E, N), color = "darkgreen") +
  ggtitle("TrajID: 4")

tra5 <- ggplot() +
  geom_point(pedestrian, mapping = aes(E, N), alpha = 0.1) +
  geom_point(subset(pedestrian, TrajID %in% "5"), mapping = aes(E, N), color = "blue") +
  geom_path(subset(pedestrian, TrajID %in% "5"), mapping = aes(E, N), color = "blue") +
  ggtitle("TrajID: 5")

tra6 <- ggplot() +
  geom_point(pedestrian, mapping = aes(E, N), alpha = 0.1) +
  geom_point(subset(pedestrian, TrajID %in% "6"), mapping = aes(E, N), color = "pink") +
  geom_path(subset(pedestrian, TrajID %in% "6"), mapping = aes(E, N), color = "pink") +
  ggtitle("TrajID: 6")

grid.arrange(tra1, tra2, tra3, tra4, tra5, tra6, nrow = 2, top = "Visual Comparisons of the 6 trajectories")

```

# Task 6

```{r}

traj_1 <- subset(pedestrian, TrajID %in% 1) 
traj_1$TrajID <- NULL
traj_1$DatetimeUTC <- NULL
traj_1 <- as.matrix(traj_1)

traj_2 <- subset(pedestrian, TrajID %in% 2)
traj_2$TrajID <- NULL
traj_2$DatetimeUTC <- NULL
traj_2 <- as.matrix(traj_2)

traj_3 <- subset(pedestrian, TrajID %in% 3)
traj_3$TrajID <- NULL
traj_3$DatetimeUTC <- NULL
traj_3 <- as.matrix(traj_3)

traj_4 <- subset(pedestrian, TrajID %in% 4)
traj_4$TrajID <- NULL
traj_4$DatetimeUTC <- NULL
traj_4 <- as.matrix(traj_4)

traj_5 <- subset(pedestrian, TrajID %in% 5)
traj_5$TrajID <- NULL
traj_5$DatetimeUTC <- NULL
traj_5 <- as.matrix(traj_5)

traj_6 <- subset(pedestrian, TrajID %in% 6)
traj_6$TrajID <- NULL
traj_6$DatetimeUTC <- NULL
traj_6 <- as.matrix(traj_6)

dtw_12 <- DTW(traj_1, traj_2, 4)
dtw_13 <- DTW(traj_1, traj_3, 4)
dtw_14 <- DTW(traj_1, traj_4, 4)
dtw_15 <- DTW(traj_1, traj_5, 4)
dtw_16 <- DTW(traj_1, traj_6, 4)

edit_12 <- EditDist(traj_1, traj_2, 20)
edit_13 <- EditDist(traj_1, traj_3, 20)
edit_14 <- EditDist(traj_1, traj_4, 20)
edit_15 <- EditDist(traj_1, traj_5, 20)
edit_16 <- EditDist(traj_1, traj_6, 20)

fre_12 <- Frechet(traj_1, traj_2)
fre_13 <- Frechet(traj_1, traj_3)
fre_14 <- Frechet(traj_1, traj_4)
fre_15 <- Frechet(traj_1, traj_5)
fre_16 <- Frechet(traj_1, traj_6)

dtws <- data.frame(traj2 = c("2", "3", "4", "5", "6"), value = c(dtw_12, dtw_13, dtw_14, dtw_15, dtw_16))

edits <- data.frame(traj2 = c("2", "3", "4", "5", "6"), value = c(edit_12, edit_13, edit_14, edit_15, edit_16))

fres <- data.frame(traj2 = c("2", "3", "4", "5", "6"), value = c(fre_12, fre_13, fre_14, fre_15, fre_16))

plot_dtw <- ggplot() +
  geom_bar(dtws, mapping = aes(traj2, value, fill = traj2), stat = "identity") +
  theme(legend.position="none") +
  ggtitle("DTW")

plot_edits <- ggplot() +
  geom_bar(edits, mapping = aes(traj2, value, fill = traj2), stat = "identity")+
  theme(legend.position="none") +
  ggtitle("EditDist")

plot_fres <- ggplot() +
  geom_bar(fres, mapping = aes(traj2, value, fill = traj2), stat = "identity") +
  theme(legend.position="none") +
  ggtitle("Frechet")

grid.arrange(plot_dtw, plot_edits, plot_fres, nrow = 2, top = "Computed similarities using different measures between trajectory 1 to all other trajectories")


```
The connections of the properties to the computed values is kind of hard to spot by eye, as most of the trajectories look the same. Trajectory 4 has a visually distinct trajectory which is kind of represented on a relatively low DTW and a medium Frechet value, but then again a much more similar (by eye) trajectory like 6 does have much smaller values.There seems to be a problem with the similarity computation sometimes (like DTV between 1 and 5 or Frechet between 1 and 2) but I couldn't make out the problem. 
